Pre working group data exploration: Mainly wrt coral cover

### 2024-07-29 julianna_data_exploration.Rmd
Fixed potential "fig 1" to be by site instead of island side and include lines for each of the disturbances in Lauren's Excel

@@

quad4sum %>% 
  # summarize and group quadrats by island/habitat
  group_by(year, habitat, island_side, substrate) %>% 
  summarise(mean_cov = mean(sum_cover),
            sd_cov = sd(sum_cover, na.rm = TRUE),
            n = n(),
            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
  # relevel for graphing
  mutate(island_side = factor(island_side, levels = c("north", "east", "west")),
         habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) -> quadHabIS

@@

```{r plot cover through time}
quadHabIS %>% 
  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
  geom_ribbon(aes(ymin = mean_cov -CI95, 
                    ymax = mean_cov + CI95),
              alpha  = 0.7) +
  geom_line() +
  theme_bw() +
  scale_fill_manual(values = c(
    moma.colors("Klein", type = "discrete")[2], # CCA
    moma.colors("Klein", type = "discrete")[1], # coral
    moma.colors("Klein", type = "discrete")[5], # macro
    moma.colors("Klein", type = "discrete")[4] # turf
  )) +
  facet_grid(island_side ~ habitat) +
  ylab("Mean cover (%)") +
  xlab("Year")
```

>>>>

# get ranges to calculate over
startAlg <- grep("acanthophora_spicifera", colnames(algae))
endAlg <- grep("valonia_ventricosa", colnames(algae))

@@

quad4sum %>% 
  # summarize and group quadrats by island/habitat
  group_by(year, habitat, island_side, substrate) %>% 
  summarise(mean_cov = mean(sum_cover),
            sd_cov = sd(sum_cover, na.rm = TRUE),
            n = n(),
            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
  # relevel for graphing
  mutate(island_side = str_to_sentence(island_side),
         island_side = factor(island_side, levels = c("North", "East", "West")),
         habitat = case_when(
           str_detect(habitat, pattern = "_") ~ str_to_sentence(paste0(str_split(habitat, "_")[[1]][1], " ", str_split(habitat, "_")[[1]][2])),
           TRUE ~ str_to_sentence(habitat)),
         habitat = factor(habitat, levels = c("Fringing", "Backreef", "Outer 10", "Outer 17"))) -> quadHabIS

@@

```{r plot cover through time}
quadHabIS %>% 
  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
  geom_ribbon(aes(ymin = mean_cov -CI95, 
                    ymax = mean_cov + CI95),
              alpha  = 0.7) +
  geom_line() +
  theme_bw() +
  scale_fill_manual(values = c(
    moma.colors("Klein", type = "discrete")[2], # CCA
    moma.colors("Klein", type = "discrete")[1], # coral
    moma.colors("Klein", type = "discrete")[5], # macro
    moma.colors("Klein", type = "discrete")[4] # turf
  ),
  labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
  facet_grid(island_side ~ habitat) +
  ylab("Mean cover (%)") +
  xlab("Year")
```

## Quarat-level habitat-site

```{r get summary}

quad4sum %>% 
  # summarize and group quadrats by island/habitat
  group_by(year, habitat, site, island_side, substrate) %>% 
  summarise(mean_cov = mean(sum_cover),
            sd_cov = sd(sum_cover, na.rm = TRUE),
            n = n(),
            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
  # relevel for graphing
  mutate(island_side = str_to_sentence(island_side),
         island_side = factor(island_side, levels = c("North", "East", "West")),
         habitat = case_when(
           str_detect(habitat, pattern = "_") ~ str_to_sentence(paste0(str_split(habitat, "_")[[1]][1], " ", str_split(habitat, "_")[[1]][2])),
           TRUE ~ str_to_sentence(habitat)),
         habitat = factor(habitat, levels = c("Fringing", "Backreef", "Outer 10", "Outer 17"))) %>% 
  mutate(site = case_when(
    site == "lter_1" ~ "LTER 1 (N)",
    site == "lter_2" ~ "LTER 2 (N)",
    site == "lter_3" ~ "LTER 3 (E)",
    site == "lter_4" ~ "LTER 4 (E)",
    site == "lter_5" ~ "LTER 5 (W)",
    site == "lter_6" ~ "LTER 6 (W)"
  )) -> quadHabSite

```{r plot cover through time}
# pdf("Fig1_quadHabSiteNo20.pdf", height = 6.5, width = 12)

quadHabSite %>% 
  filter(year != 2020) %>% 
  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
  geom_ribbon(aes(ymin = mean_cov -CI95, 
                    ymax = mean_cov + CI95),
              alpha  = 0.7) +
  geom_line() +
  theme_bw() +
  scale_fill_manual(values = c(
    moma.colors("Klein", type = "discrete")[2], # CCA
    moma.colors("Klein", type = "discrete")[1], # coral
    moma.colors("Klein", type = "discrete")[5], # macro
    moma.colors("Klein", type = "discrete")[4] # turf
  ),
  labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
  facet_grid(habitat ~ site) + # not sure which direction is best here
  ylab("Mean cover (%)") +
  xlab("Year") +
  # add lines for disturbances
  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
  geom_vline(xintercept = 2019, linetype = "dashed", color = "red")  # bleaching

# dev.off()
```
